HyperLogLog: Exponentially Bad in Adversarial Settings

### 2020-10-04 docs/HLL_exp_bad/main.tex
some more notes and tweaks

\subsubsection{Initially Non-empty HLL Sketch, Scenario S4:}\label{sec:nonemptyHLLS4}
The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch). Therefore, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate. In other words, if $h(x)$ is mapped to the $j$-th bucket containing $M[j] \ge 1$, then, instead of the strict requirement of having a ``1'' bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a ``1'' bit in any of the first $M[j]$ positions in this substring. Under these conditions, the estimates in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.

@@

\todo{Fix this -- it's not clear the attack works because it's not clear we really can detect cardinality changes in the way desired (and we've not demonstrated that we can).}

@@

>>>>

\subsubsection{Initially Non-empty HLL Sketch, Scenario S4:}\label{sec:nonemptyHLLS4}
The adversary can adopt the same strategy as for S2, but make relax the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch). Specifically, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate. In other words, if $h(x)$ is mapped to the $j$-th bucket containing $M[j] \ge 1$, then, instead of the strict requirement of having a ``1'' bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a ``1'' bit in any of the first $M[j]$ positions in this substring. Under these conditions, the estimates in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.

@@

\todo{Fix this -- it's not clear the attack works because it's not clear we really can detect cardinality changes in the way desired (and we've not demonstrated that we can). Explain why it does not work -- to stop reviewers finding wrong solutions.}

@@

\todo{include some urls for code and release notes.}
